---
title: "PA1_template"
author: "Javier"
date: "31 de octubre de 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Reproducible Research: Peer Assigment 1

## Loading and Processing Data
If you already download the zip file, please extract the files in your working directory and keep the names of file as "activity.csv".
If you haven´t, the code will download the zip file and extract the files for you.
```{r}
WD <- getwd()
dest_file1 <- paste(WD, "/activity.csv")
dest_file1 <- gsub(" /", "/", dest_file1)

if (!file.exists(dest_file1, showWarnings = FALSE)){
  URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  temp <- tempfile()
  download.file(URL, temp)
  unzip(temp, c("activity.csv"), exdir = WD )
  ActivityData <- read.csv("activity.csv")
  unlink(temp)
}else{
  ActivityData <- read.csv("activity.csv")
}
```

## What is mean total number of steps taken per day?
First we calculate the total number of steps taken per day:
```{r}
library(ggplot2)
TotalSteps <- tapply(ActivityData$steps, ActivityData$date, FUN = sum, na.rm = TRUE)
```
Then, we make a histogram of the total number of steps taken each day:
```{r}
qplot(TotalSteps, binwidth = 1000, xlab = "Total number of steps taken each day")
```
Finally, we calculate and report the mean and median of the total number of steps taken per day:
```{r}
Mean1 <- mean(TotalSteps, na.rm = TRUE)
Median1 <- median(TotalSteps, na.rm = TRUE)
```

## What is the average daily activity pattern?
First, we make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r}
Averages <- aggregate(x=list(steps = ActivityData$steps), by=list(interval = ActivityData$interval), FUN = mean, na.rm = TRUE)
ggplot(data = Averages, aes(x=interval, y=steps)) + geom_line() + xlab("5 minute interval") + ylab("average number of steps taken")
```

Finally, we calculate the maximum number of steps on average across all the days in the dataset in a 5-minutes interval:
```{r}
Averages[which.max(Averages$steps),]
```

## Imputing missing values
we calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs) as follows:
```{r}
Missing <- is.na(ActivityData$steps)
table(Missing)
```

Then, all the missing values are filled in with mean value for that 5-minute interval.
```{r}
FillValue <- function(steps, interval) {
  Filled <- NA
  if (!is.na(steps))
    Filled <- c(steps)
  else
    Filled <- (Averages[Averages$interval==interval, "steps"])
  return(Filled)
}
FilledData <- ActivityData
FilledData$steps <- mapply(FillValue, FilledData$steps, FilledData$interval)
```

Finally, we used the filled data set to make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.
```{r}
library(ggplot2)
TotalSteps <- tapply(ActivityData$steps, ActivityData$date, FUN = sum, na.rm = TRUE)
qplot(TotalSteps, binwidth = 1000, xlab = "Total number of steps taken each day")
```

```{r}
Mean2 <- mean(TotalSteps, na.rm = TRUE)
Median2 <- median(TotalSteps, na.rm = TRUE)
```
Do these values differ from the estimates from the first part of the assignment?
In the first part of the assigment, the obtained mean was `Mean1` and the obtained median was `Median1`. 

What is the impact of imputing missing data on the estimates of the total daily number of steps?
As you can notice, the mean `Mean2` and `Median2` are higher because we replaced the missing values with the mean associated interval values. This replacement has a impact on the data because the NA values are now real data related to the mean values.

## Are there differences in activity patterns between weekdays and weekends?
First, we create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
WeekdayOrWeekend <- function(Date) {
  Day <- weekdays(Date)
  if (Day %in% c("lunes", "martes", "miércoles", "jueves", "viernes"))
    return("Weekday")
  else if (Day %in% c("sábado", "domingo"))
    return("Weekend")
  else
    stop("Invalid date")
}
FilledData$date <- as.Date(FilledData$date)
FilledData$day <- sapply(FilledData$date, FUN = WeekdayOrWeekend)
```

Finally, we make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
Averages <- aggregate(steps ~ interval + day, data = FilledData, mean)
ggplot(Averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
 xlab("5-minute interval") + ylab("Number of steps")
```


